---
title: "CobbDouglas"
author: "Dr. J"
date: "4/24/2020"
output: html_document
---

## non-linear optimization

This Rmarkdown file demonstrates techniques for non-linear optimization, and is illustrated with data for the Cobb-Douglas production function, $y = A K^\alpha L^\beta$, K capital, L labor, and the exponents the elasticies of production by input.

### Methods demonstrated:
 1. Grid search of space of parameters  
 2. non-linear optimization software, find values of the parameters that minimize the sum of squared errors, errors are (data-predicted)  
 3. non-linear optimizatin software, maximum likelihood- find values of the parameters that maximize the likelihood of the data assuming errors are normally distributed  
 4. log-transform to a linear model  
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data

```{r}
CobbDoug = read.csv('CobbDouglasData.csv')
attach(CobbDoug)
```


```{r}
plot(Cap, Q, pch=19, col='red', xlab= 'input values', ylab = 'output',
     main='Cobb-Douglas Production function')
points(Emp, Q, pch=19, col='blue')
grid()
```

A two-dimensional plot would be nice..

```{r}

```

Set up a non-linear loss function

```{r}
CDloss=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  Loss = 0
  for (i in 1:length(Q)){
    Loss = Loss + ((Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))^2)
  }
  
  return (log(Loss))
}

CDmle=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  logLoss = 0
  for (i in 1:length(Q)){
    errorTerm= (Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))
    logLoss = logLoss - dnorm(errorTerm, log=T)
  }
  
  return (logLoss)
}
```

Test the loss function

```{r}
A = 1
alpha = 0.5
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))

# better or worse than
A = 0.5
alpha = 0.2
beta = 0.8
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))
```
## Grid search
  not the best idea, can get close to an answer 

```{r}

for (A in seq(from=0.01, to = 2, length.out = 4))
  for(a in seq(from=0.01, to = 1,length.out = 6))
   for(b in seq(from=0.01, to = 1, length.out = 6)){
     parms = c(A, a, b)
      print(paste(A, a, b,'        ', CDloss(parms, Q, Cap, Emp)))
  }

```

## non linear optimization

use function optim to find values of A, alpha, beta
find values that minimize the loss function, loss = sum of sqrared errors.  This is a 'least squares' solution

```{r}
##     ?optim
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDloss, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDloss(parms, Q, Cap,Emp))
```

## maximum likelihood estimation

With the right fit, the residual errors should be independent, normally distributed, with mean 0. likelihood of the parameter set (a, $\alpha$, $\beta$) is

$\prod_0^n \phi(e_i)$

$\phi$ is the standard normal distribution


```{r}
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDmle, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDloss(parms, Q, Cap,Emp))
```



## transform to logs

```{r}
logModel = glm(log(Q)~log(Cap)+log(Emp))
summary(logModel)
A = exp(logModel$coefficients[1])
alpha = logModel$coefficients[2]
beta = logModel$coefficients[3]
print(paste('A = ',A,' alpha = ', alpha,' beta =', beta))

# CD loss at these values:
parms=c(A, alpha, beta)
print(paste('CD loss: ', CDloss(parms, Q, Cap, Emp)))
```

## Questions:
1 what are the estimated values of A, alpha and beta
2 confidence intervals on A, alpha, beta?
3 is production at constant return to scale? If production is increasing or decreasing return to scale, which input is most expensive
4. What does the values of A mean??

