---
title: "CobbDouglas"
author: "Dr. J"
date: "4/24/2020"
output: html_document
---


This Rmarkdown file demonstrates techniques for non-linear optimization. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data

```{r}
CobbDoug = read.csv('CobbDouglasData.csv')
attach(CobbDoug)
```


```{r}
plot(Cap, Q, pch=19, col='red', xlab= 'input values', ylab = 'output',
     main='Cobb-Douglas Production function')
points(Emp, Q, pch=19, col='blue')
grid()
```

A two-dimensional plot would be nice..

```{r}

```

Set up a non-linear loss function

```{r}
CDloss=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  logLoss = 0
  for (i in 1:length(Q)){
    logLoss = logLoss + log((Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))^2)
  }
  
  return (logLoss)
}
```

Test the loss function

```{r}
A = 1
alpha = 0.5
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))

# better or worse than
A = 0.5
alpha = 0.2
beta = 0.8
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))
```
## Echaustive search
  not the best idea, can get close to an answer 

```{r}

for (A in seq(from=0.1, to = 3, length.out = 5))
  for(a in seq(from=0.1, to = 1,length.out = 5))
   for(b in seq(from=0.1, to = 1, length.out = 5)){
     parms = c(A, a, b)
      print(paste(A, a, b,'        ', CDloss(parms, Q, Cap, Emp)))
  }

```


## non linear optimization
use function optim to find values of A, alpha, beta
find values that minimize the loss function, loss = sum of sqrared errors

```{r}
##     ?optim
parms = c(1, 1.15, 10.15)
opt = optim(parms, CDloss, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
```

## transform to logs

```{r}
logModel = glm(log(Q)~log(Cap)+log(Emp))
summary(logModel)
A = exp(logModel$coefficients[1])
alpha = logModel$coefficients[2]
beta = logModel$coefficients[3]
print(paste('A = ',A,' alpha = ', alpha,' beta =', beta))

# CD loss at these values:
parms=c(A, alpha, beta)
print(paste('CD loss: ', CDloss(parms, Q, Cap, Emp)))
```

## Questions:
1 what are the estimated values of A, alpha and beta
2 confidence intervals on alpha, beta?
3 is production at constant return to scale?
4. What does the values of A mean??

